---
title: "R Notebook"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
numbersections: yes
---


# Statistical analysis: IGN game review

### Introduction

This R notebook uses the IGN game dataset.

The goals/questions to answer are:

* How do the sales relate to the scores, platform, genre or platform type?
* Do the scores from IGN relate to the scores from other sources? For example, reviews from users collected from GameSpot.
* What are the differences between the consoles from the same time period? For example, the consoles at the time period of PlayStation 2 vs. XBOX, PlayStation 1 vs Nintendo 64...

I will perform some statistical analysis (Anderson-Darling normality test, Wilcoxon rank sum test with continuity correction, mean via Bootstrapping) and visualizations to answer the previous set of questions.

### Data collection and Data cleaning

I use the games scores and sales from VGChartz http://www.vgchartz.com which contains a compilation of scores from different webpages and critics for the last $\sim$40 years. The data has been obtained by scrapping the webpage and stored in a csv file.

With python we have cleaned the data. In short, we performed the following:

+ We removed the fields with empty game name (unique)
+ We replaced the empty cells by NaNs
+ We created a new variable which groups the different platforms into three categories: Console, Handheld System and OS.

```{r}
library(ggplot2)
library(gridExtra)
library(pwr)
library(gtools)
library(nortest)
library(boot)
library(plyr)
```

```{r}
df=read.csv("VG_final.csv")
str(df)
```


### Data analysis and modelling

We want to understand whether the game scores vary from year to year. We see how the scores vary as a function of the release year.

```{r}
pairs(~Year_of_Release+Critic_Score+User_Score,data=df)
```

We see that before 1996 there are not enough scores After that, it is really hard to see a variation of the scores with the year. We see that critic score and user score are related.

```{r}
pairs(~Year_of_Release[Year_of_Release>1996]+Critic_Score[Year_of_Release>1996]+User_Score[Year_of_Release>1996],data=df)
```

To see whether our variables show a normal distribution, we perform the 

```{r}
ad.test(df[df$Year_of_Release>1996,]$Year_of_Release)
```

The distribution is not normal, so we perform the non-parametric Wilcox test:

```{r}
wilcox.test(df[df$Year_of_Release>1996,]$Year_of_Release,df[df$Year_of_Release>1996,]$User_Score, paired=FALSE, alternative='greater')
```
So their means are not correlated.

We see how the mean scores varies over years. For that, we do bootstrapping to compute the mean scores.

```{r}
# function to compute the mean through bootstrap
boot_mean <- function(data){
samples <- replicate(length(data), mean(sample(data, size = length(data), replace = TRUE)))
  return(samples)
}
```





```{r}
means_year=NULL
stds_year=NULL
for (i in 1:length(unique(df$Year_of_Release))){
  if (length(na.exclude(df[df$Year_of_Release==unique(df$Year_of_Release)[i],]$User_Score))>3){
      means_year_distr=boot_mean(na.exclude(df[df$Year_of_Release==unique(df$Year_of_Release)[i],]$User_Score))
     means_year[i]=mean(means_year_distr, na.rm = TRUE)
     stds_year[i]=sd(means_year_distr, na.rm = TRUE)
  } else{
    means_year[i]=NA
    stds_year[i]=NA
  }
}
df_means_year=data.frame("years"=unique(df$Year_of_Release),"Means"=means_year,"Std"=stds_year)
plot1= ggplot(df_means_year, aes(x = years,y=Means),color='blue') +
  geom_errorbar(aes(ymin=Means-stds_year,ymax=Means+stds_year,color=Means))+labs(title = "Averages of User Scores", x = "Year", y = "Average score")+xlim(1995,2017)+ylim(6,9) +    geom_line()+geom_hline(aes(yintercept = mean(df_means_year$Means,na.rm = TRUE)),color='red',linetype="dashed", size = .4)
```



We plot the results:


```{r}
means_year=NULL
stds_year=NULL
for (i in 1:length(unique(df$Year_of_Release))){
  if (length(na.exclude(df[df$Year_of_Release==unique(df$Year_of_Release)[i],]$Critic_Score))>3){
      means_year_distr=boot_mean(na.exclude(df[df$Year_of_Release==unique(df$Year_of_Release)[i],]$Critic_Score))
     means_year[i]=mean(means_year_distr, na.rm = TRUE)
     stds_year[i]=sd(means_year_distr, na.rm = TRUE)
  } else{
    means_year[i]=NA
    stds_year[i]=NA
  }
}
df_means_year2=data.frame("years"=unique(df$Year_of_Release),"Means"=means_year,"Std"=stds_year)
plot2= ggplot(df_means_year2, aes(x = years,y=Means),color='blue') +
  geom_errorbar(aes(ymin=Means-stds_year,ymax=Means+stds_year,color=Means))+labs(title = "Averages of Critic Scores", x = "Year", y = "Average score") +    geom_line()+geom_hline(aes(yintercept = mean(df_means_year2$Means,na.rm = TRUE)),color='red',linetype="dashed", size = .4)+xlim(1995,2017)+ylim(60,90)

```

```{r fig2, fig.height = 4, fig.width = 10}
grid.arrange(plot1,plot2,ncol=2)
```


### Which platform has the highest average reviews?

We compare the average scores from both users and critics, and see which platform got the better reviews.

```{r}
means_plat=NULL
stds_plat=NULL
for (i in 1:length(levels(df$Platform))){
  if (length(na.exclude(df[df$Platform==levels(df$Platform)[i],]$User_Score))>30){
    means_plat_distr=boot_mean(na.exclude(df[df$Platform==levels(df$Platform)[i],]$User_Score))
    means_plat[i]=mean(means_plat_distr, na.rm = TRUE)
    stds_plat[i]=sd(means_plat_distr, na.rm = TRUE)
    } else{
    means_plat[i]=NA
    stds_plat[i]=NA
  }
}
df_means_plat=data.frame("Platform"=levels(df$Platform),"Means"=means_plat,"Std"=stds_plat)
df_means_plat2=subset(df_means_plat,!(is.na(df_means_plat["Std"])))
df_means_plat3=df_means_plat2[order(df_means_plat2$Means),]
mean_of_means=('Average_of_means'=mean(df_means_plat2$Means))
plot3= ggplot(df_means_plat3, aes(x = Platform,y=Means)) +
  geom_errorbar(aes(ymin=Means-Std,ymax=Means+Std,color=Means))+labs(title = "Averages of User Scores per Platform", x = "Platform", y = "Average score") + geom_line() +geom_hline(aes(yintercept = mean_of_means),color='red',linetype="dashed", size = .4)+ylim(6,8)
```



```{r}
means_plat=NULL
stds_plat=NULL
for (i in 1:length(levels(df$Platform))){
  if (length(na.exclude(df[df$Platform==levels(df$Platform)[i],]$Critic_Score))>30){
    means_plat_distr=boot_mean(na.exclude(df[df$Platform==levels(df$Platform)[i],]$Critic_Score))
    means_plat[i]=mean(means_plat_distr, na.rm = TRUE)
    stds_plat[i]=sd(means_plat_distr, na.rm = TRUE)
    } else{
    means_plat[i]=NA
    stds_plat[i]=NA
  }
}
df_means_plat4=data.frame("Platform"=levels(df$Platform),"Means"=means_plat,"Std"=stds_plat)
df_means_plat5=subset(df_means_plat4,!(is.na(df_means_plat4["Std"])))
df_means_plat6=df_means_plat5[order(df_means_plat5$Means),]
mean_of_means2=('Average_of_means'=mean(df_means_plat6$Means))
plot4= ggplot(df_means_plat6, aes(x = Platform,y=Means)) +
  geom_errorbar(aes(ymin=Means-Std,ymax=Means+Std,color=Means))+labs(title = "Averages of Critic Scores per Platform", x = "Platform", y = "Average score") + geom_line() +geom_hline(aes(yintercept = mean_of_means2),color='red',linetype="dashed", size = .4)+ylim(60,80)
```

```{r fig1, fig.height = 8, fig.width = 10}
grid.arrange(plot3,plot4,ncol=1)
```




